---
title: "Supplementary Code 1"
output: html_document
date: "2025-03-18"
---

```{r}
library(dplyr)
library(tidyverse)
```

```{r}
#Load dataset (the mammals categorized, the plants they disperse, and the plant traits)
systematic_review_analysis <- read.csv("systematic_review_analysis.csv")

#Load lists of trait values for each plant species
all_native_introduced <- read.csv("all_native_introduced.csv")
all_dehiscent_indehiscent <- read.csv("all_dehiscent_indehiscent.csv")
all_fleshy_dry <- read.csv("all_fleshy_dry.csv")
all_num_seeds <- read.csv("all_num_seeds.csv")
all_s_width <- read.csv("all_s_width.csv")
all_f_width <- read.csv("all_f_width.csv")
```

Permutation test for discrete traits (native/introduced, dehiscent/indehiscent, fleshy/dry)

```{r}
permutation_discrete <- function(plant_trait,plant_df,trait_value_1,trait_value_2,mammal_trait,mammal_extant){
  #Calculate test statistic (proportional change) of original data:
  #Calculate proportional change for trait value 1
  filter_detail_1 <- systematic_review_analysis %>% filter({{plant_trait}} == trait_value_1) #Keep only the rows with plants with trait value 1 
  before_extinction_1 <- nrow(filter_detail_1) #Calculate the number of mammal-plant links with trait value 1
  extinction_filter_1 <- filter_detail_1 %>% filter({{mammal_trait}} == mammal_extant) #Keep only the rows with non-threatened or small-bodied mammals
  after_extinction_1 <- nrow(extinction_filter_1) #Count the number of rows with non-threatened or small-bodied mammals interacting with plants with trait value 1
  change_1 <- abs((before_extinction_1 - after_extinction_1) / before_extinction_1)

  #Calculate proportional change for trait value 2
  filter_detail_2 <- systematic_review_analysis %>% filter({{plant_trait}} == trait_value_2) 
  before_extinction_2 <- nrow(filter_detail_2) 
  extinction_filter_2 <- filter_detail_2 %>% filter({{mammal_trait}} == mammal_extant)
  after_extinction_2 <- nrow(extinction_filter_2) 
  change_2 <- abs((before_extinction_2 - after_extinction_2) / before_extinction_2)
  
  #Calculate test statistic
  true_test_statistic <- change_1-change_2
  print(true_test_statistic)
  
  #Create a dataframe with all the possible links between mammals and plants on which we have data for a given plant trait
  systematic_review_clean <- filter(systematic_review_analysis,!is.na({{plant_trait}})) #Remove all plant species for which we don't have a measurement
  mammals_unique <- length(unique(systematic_review_clean$Mammal)) #Find the number of mammals in the dataset for that plant trait
  plants_duplicate <- cbind(plant_df,rep(row.names(plant_df),each = mammals_unique)) #Create a dataframe where all of the plants are repeated for each mammal species 
  mammals <- systematic_review_clean[,c(1:3)] #Separate the columns containing the mammal, threatened status, and body size
  mammals <- distinct(mammals) #Remove duplicate rows so there is only one row per mammal species
  mammals_duplicate <- cbind(mammals,rep(row.names(mammals),each=length(plant_df$Plant_Species))) #Create a dataframe where all the mammals are repeated for each plant species
  mammals_duplicate <- arrange(mammals_duplicate,Mammal) #Put the dataframe in alphabetical order
  all <- cbind(mammals_duplicate,plants_duplicate) #Combine mammal and plant lists into one dataframe with all possible interactions between mammals and plants
  
  #Sample from the dataframe of all possibilities the number of links each mammal has in the network. Repeat this for each permutation.
  link_counts <- group_by(systematic_review_clean,Mammal) %>% summarize(Count = n()) #Count the number of links per mammal species in the original dataset
  mammal_links <- with(link_counts,split(Count,factor(Mammal,levels=unique(Mammal)))) #Turn link_counts into a list
  permutation_test_statistics <- numeric(10000) #Create an empty vector to store the test statistics for the 10,000 permuted datasets
  
  set.seed(100)
  for (i in 1:10000) {
  mammal_groups <- split(all, all$Mammal) #Split dataframe with all possibilities up by mammal species
  
  #For each mammal, remove links until the number of links is the same as in the original dataset
  for (Mammal in names(mammal_groups)) {
    links_current <- mammal_links[[Mammal]] #Number of links in the original dataset
    if (nrow(mammal_groups[[Mammal]]) > links_current) {
      sample_mammal <- sample(nrow(mammal_groups[[Mammal]]), links_current) #Randomly sample indices until read the number in the original dataset for that mammal
      mammal_groups[[Mammal]] <- mammal_groups[[Mammal]][sample_mammal, ]
    }
  }
  
  final_permutation <- do.call(rbind, mammal_groups) #Combine outputs for each mammal maintaining set order in df

  #Calculate the test statistic for each permuted dataset:
  #Calculate proportional change for trait value 1
  filter_detail_1 <- final_permutation %>% filter({{plant_trait}} == trait_value_1) #Keep only the rows with plants with trait value 1 
  before_extinction_1 <- nrow(filter_detail_1) #Calculate the number of mammal-plant links with trait value 1
  extinction_filter_1 <- filter_detail_1 %>% filter({{mammal_trait}} == mammal_extant) #Keep only the rows with non-threatened or small-bodied mammals
  after_extinction_1 <- nrow(extinction_filter_1) #Count the number of rows with non-threatened or small-bodied mammals interacting with plants with trait value 1
  change_1 <- abs((before_extinction_1 - after_extinction_1) / before_extinction_1)

  #Calculate proportional change for trait value 2
  filter_detail_2 <- final_permutation %>% filter({{plant_trait}} == trait_value_2) 
  before_extinction_2 <- nrow(filter_detail_2) 
  extinction_filter_2 <- filter_detail_2 %>% filter({{mammal_trait}} == mammal_extant)
  after_extinction_2 <- nrow(extinction_filter_2) 
  change_2 <- abs((before_extinction_2 - after_extinction_2) / before_extinction_2)
  
  #Calculate test statistic and store in permutation_test_statistics vector
  test_statistic <- change_1-change_2
  permutation_test_statistics[i] <- test_statistic
  }

#Run a two-tailed test 
two_tailed <- sum(permutation_test_statistics>abs(true_test_statistic))+sum(permutation_test_statistics<(-abs(true_test_statistic))) #Proportion of results in the two tails
p_value <- two_tailed/10000
print(p_value)
}
```

Example of implementation for testing the effect large-bodied mammal extinction on fleshy vs. dry fruit

```{r}
permutation_discrete(Fleshy_Dry,all_fleshy_dry,"F","D",Body_Size,"S")
```

Permutation test for continuous traits (number of seeds/fruit, seed width, fruit width)

```{r}
permutation_continuous <- function(plant_trait,plant_df,mammal_trait,mammal_extant){
  #Calculate test statistic (mean pre-extinction - mean post-extinction) of original data:
  systematic_review_clean <- filter(systematic_review_analysis,!is.na({{plant_trait}})) #Remove all plant species for which we don't have a measurement
  before_extinction <- systematic_review_clean %>% summarise(mean_value=mean({{plant_trait}})) %>% pull(mean_value) #Calculate the mean plant trait dispersed before extinction
  extinction_filter <- systematic_review_clean %>% filter({{mammal_trait}} == mammal_extant) #Keep only the rows with non-threatened or small-bodied mammals
  after_extinction <- extinction_filter %>% summarise(mean_value=mean({{plant_trait}})) %>% pull(mean_value) #Calculate the mean plant trait dispersed after extinction of large-bodied mammals
  true_test_statistic <- before_extinction - after_extinction
  print(true_test_statistic)

  #Create a dataframe with all the possible links between mammals and plants on which we have data for a given plant trait
  mammals_unique <- length(unique(systematic_review_clean$Mammal)) #Find the number of mammals in the dataset for that plant trait
  plants_duplicate <- cbind(plant_df,rep(row.names(plant_df),each = mammals_unique)) #Create a dataframe where all of the plants are repeated for each mammal species 
  mammals <- systematic_review_clean[,c(1:3)] #Separate the columns containing the mammal, threatened status, and body size
  mammals <- distinct(mammals) #Remove duplicate rows so there is only one row per mammal species
  mammals_duplicate <- cbind(mammals,rep(row.names(mammals),each=length(plant_df$Plant_Species))) #Create a dataframe where all the mammals are repeated for each plant species
  mammals_duplicate <- arrange(mammals_duplicate,Mammal) #Put the dataframe in alphabetical order
  all <- cbind(mammals_duplicate,plants_duplicate) #Combine mammal and plant lists into one dataframe with all possible interactions between mammals and plants
  
  #Sample from the dataframe of all possibilities the number of links each mammal has in the network. Repeat this for each permutation.
  link_counts <- group_by(systematic_review_clean,Mammal) %>% summarize(Count = n()) #Count the number of links per mammal species in the original dataset
  mammal_links <- with(link_counts,split(Count,factor(Mammal,levels=unique(Mammal)))) #Turn link_counts into a list
  permutation_test_statistics <- numeric(10000) #Create an empty vector to store the test statistics for the 10,000 permuted datasets
  
  set.seed(100)
  for (i in 1:10000) {
  mammal_groups <- split(all, all$Mammal) #Split dataframe with all possibilities up by mammal species
  
  #For each mammal, remove links until the number of links is the same as in the original dataset
  for (Mammal in names(mammal_groups)) {
    links_current <- mammal_links[[Mammal]] #Number of links in the original dataset
    if (nrow(mammal_groups[[Mammal]]) > links_current) {
      sample_mammal <- sample(nrow(mammal_groups[[Mammal]]), links_current) #Randomly sample indices until read the number in the original dataset for that mammal
      mammal_groups[[Mammal]] <- mammal_groups[[Mammal]][sample_mammal, ]
    }
  }
  
  final_permutation <- do.call(rbind, mammal_groups) #Combine outputs for each mammal maintaining set order in df

  #Calculate the test statistic for each permuted dataset:
  before_extinction <- final_permutation %>% summarise(mean_value=mean({{plant_trait}})) %>% pull(mean_value) #Calculate the mean plant trait dispersed before extinction
  extinction_filter <- final_permutation %>% filter({{mammal_trait}} == mammal_extant) #Keep only the rows with non-threatened or small-bodied mammals
  after_extinction <- extinction_filter %>% summarise(mean_value=mean({{plant_trait}})) %>% pull(mean_value) #Calculate the mean plant trait dispersed after extinction of large-bodied mammals
  test_statistic <- before_extinction - after_extinction
  permutation_test_statistics[i] <- test_statistic
  }

#Run a two-tailed test 
two_tailed <- sum(permutation_test_statistics>abs(true_test_statistic))+sum(permutation_test_statistics<(-abs(true_test_statistic))) #Proportion of results in the two tails
p_value <- two_tailed/10000
print(p_value)
}
```

Example of implementation for testing the effect threatened mammal extinction on the number of seeds per fruit

```{r}
permutation_continuous(Number_S_per_F,all_num_seeds,Threatened_Status,"NT")
```
